---
title: "Modify NASCAR.data for machine learning"
author: "Gerald Belton"
date: "November 27, 2015"
output: html_document
---
Start by reducing NASCAR.data to just the variables we will use for the machine 
learning process:
```{r}
# Subset NASCAR.data to produce training data set
# Load existing database
load("NASCARdata.RDa")
keepvars <- c("Date", "Site", "Driver", "Start", "Rating", "Finish")
training.data <- NASCAR.data[keepvars]

#convert strings to factors
training.data$Site <- as.factor(training.data$Site)
training.data$Driver <- as.factor(training.data$Driver)
#training.data$CarNumber <- as.factor(training.data$CarNumber)

```

We're going to nead some information on track type:
```{r}
tracks <- read.csv("NASCARTracks.csv")

```

Classify each track according to NASCAR and AccuPredict classifications
```{r}
require(plyr)
training.data = join(training.data,tracks,by='Site')    
```

For each race starting with 2006, calculate each driver's previous average 
finish position.
```{r}

for (row in 1:nrow(training.data)) {
     dvr <- training.data$Driver[row]
     Dte <- training.data$Date[row]
     temp <- training.data[ which(training.data$Driver==dvr &  
                                       training.data$Date<Dte), ]
     training.data$CareerAvg[row] <- mean(temp$Finish)
     
    
}
#CareerAvg contains the driver's average finish for his career up to this race
```

Create new column Avg10 which contains the driver's average in his last ten 
starts.
```{r}


for (row in 1:nrow(training.data)) {
     dvr <- training.data$Driver[row]
     Dte <- training.data$Date[row]
     temp <- training.data[ which(training.data$Driver==dvr &  
                                       training.data$Date<Dte), ]
     training.data$Avg10[row] <- mean(tail(temp$Finish,10))
     
    
}
```

Create new column Avg5 which contains the driver's average in his last five 
starts.
```{r}


for (row in 1:nrow(training.data)) {
     dvr <- training.data$Driver[row]
     Dte <- training.data$Date[row]
     temp <- training.data[ which(training.data$Driver==dvr &  
                                       training.data$Date<Dte), ]
     training.data$Avg5[row] <- mean(tail(temp$Finish,5))
     
    
}
```

Now let's calculate the drivers average finish at the same track
```{r}

for (row in 1:nrow(training.data)) {
     dvr <- training.data$Driver[row]
     Dte <- training.data$Date[row]
     trk <- training.data$Site[row]
     temp <- training.data[ which(training.data$Driver==dvr & 
                                       training.data$Date<Dte &
                                       training.data$Site == trk), ]
     training.data$TrackAvg[row] <- mean(temp$Finish)
     
    
}
```

Calculate driver's previous average by NASCAR track type
```{r}

for (row in 1:nrow(training.data)) {
     dvr <- training.data$Driver[row]
     Dte <- training.data$Date[row]
     NSC <- training.data$NASCAR[row]
     temp <- training.data[ which(training.data$Driver==dvr & 
                                       training.data$Date<Dte &
                                       training.data$NASCAR == NSC), ]
     training.data$NTypeAvg[row] <- mean(temp$Finish)
     
    
}
```

Calculate driver's previous average by AccuPredict track type
```{r}

for (row in 1:nrow(training.data)) {
     dvr <- training.data$Driver[row]
     Dte <- training.data$Date[row]
     ACC <- training.data$AccuPredict[row]
     temp <- training.data[ which(training.data$Driver==dvr & 
                                       training.data$Date<Dte &
                                       training.data$AccuPredict == ACC), ]
     training.data$AccuPredictAvg[row] <- mean(temp$Finish)
     
    
}
```

Create new column RtgAvg10 which contains the driver's average Rating in his 
last ten starts.
```{r}


for (row in 1:nrow(training.data)) {
     dvr <- training.data$Driver[row]
     Dte <- training.data$Date[row]
     temp <- training.data[ which(training.data$Driver==dvr &  
                                       training.data$Date<Dte), ]
     training.data$RtgAvg10[row] <- mean(tail(temp$Rating,10))
     
    
}
```

Save training data for later use
```{r}
save(training.data, file="training-data.RData")
```