---
title: "Modify NASCAR.data for machine learning"
author: "Gerald Belton"
date: "November 27, 2015"
output: html_document
---
Start by reducing NASCAR.data to just the variables we will use for the machine 
learning process:
```{r}
# Subset NASCAR.data to produce training data set
# Load existing database
load("NASCARdata.RDa")
keepvars <- c("Date", "Site", "Driver", "CarNumber", "Finish")
NASCAR.data <- NASCAR.data[keepvars]

#convert strings to factors
NASCAR.data$Site <- as.factor(NASCAR.data$Site)
NASCAR.data$Driver <- as.factor(NASCAR.data$Driver)
NASCAR.data$CarNumber <- as.factor(NASCAR.data$CarNumber)

```

We're going to nead some information on track type:
```{r}
tracks <- read.csv("NASCARTracks.csv")

```

Classify each track according to NASCAR and AccuPredict classifications
```{r}
require(plyr)
NASCAR.data = join(NASCAR.data,tracks,by='Site')    
```




For each race starting with 2006, calculate each driver's previous average 
finish position.
```{r}

for (row in 1:nrow(NASCAR.data)) {
     dvr <- NASCAR.data$Driver[row]
     Dte <- NASCAR.data$Date[row]
     temp <- NASCAR.data[ which(NASCAR.data$Driver==dvr &  
                                       NASCAR.data$Date<Dte), ]
     NASCAR.data$CareerAvg[row] <- mean(temp$Finish)
     
    
}
#CareerAvg contains the driver's average finish for his career up to this race
```

Create new column Avg10 which contains the driver's average in his last ten 
starts.
```{r}


for (row in 1:nrow(NASCAR.data)) {
     dvr <- NASCAR.data$Driver[row]
     Dte <- NASCAR.data$Date[row]
     temp <- NASCAR.data[ which(NASCAR.data$Driver==dvr &  
                                       NASCAR.data$Date<Dte), ]
     NASCAR.data$Avg10[row] <- mean(tail(temp$Finish,10))
     
    
}
```

Create new column Avg5 which contains the driver's average in his last five 
starts.
```{r}


for (row in 1:nrow(NASCAR.data)) {
     dvr <- NASCAR.data$Driver[row]
     Dte <- NASCAR.data$Date[row]
     temp <- NASCAR.data[ which(NASCAR.data$Driver==dvr &  
                                       NASCAR.data$Date<Dte), ]
     NASCAR.data$Avg5[row] <- mean(tail(temp$Finish,5))
     
    
}
```

Now let's calculate the drivers average finish at the same track
```{r}

for (row in 1:nrow(NASCAR.data)) {
     dvr <- NASCAR.data$Driver[row]
     Dte <- NASCAR.data$Date[row]
     trk <- NASCAR.data$Site[row]
     temp <- NASCAR.data[ which(NASCAR.data$Driver==dvr & 
                                       NASCAR.data$Date<Dte &
                                       NASCAR.data$Site == trk), ]
     NASCAR.data$TrackAvg[row] <- mean(temp$Finish)
     
    
}
```

Calculate driver's previous average by NASCAR track type
```{r}

for (row in 1:nrow(NASCAR.data)) {
     dvr <- NASCAR.data$Driver[row]
     Dte <- NASCAR.data$Date[row]
     NSC <- NASCAR.data$NASCAR[row]
     temp <- NASCAR.data[ which(NASCAR.data$Driver==dvr & 
                                       NASCAR.data$Date<Dte &
                                       NASCAR.data$NASCAR == NSC), ]
     NASCAR.data$NTypeAvg[row] <- mean(temp$Finish)
     
    
}
```

Calculate driver's previous average by AccuPredict track type
```{r}

for (row in 1:nrow(NASCAR.data)) {
     dvr <- NASCAR.data$Driver[row]
     Dte <- NASCAR.data$Date[row]
     ACC <- NASCAR.data$AccuPredict[row]
     temp <- NASCAR.data[ which(NASCAR.data$Driver==dvr & 
                                       NASCAR.data$Date<Dte &
                                       NASCAR.data$AccuPredict == ACC), ]
     NASCAR.data$AccuPredictAvg[row] <- mean(temp$Finish)
     
    
}
```